Single-step Hybrid Predition

### 2016-12-07 reports/index.Rmd
Update equation numbering after updates to bookdown package
Describe kernel transformation in report

<script type="text/x-mathjax-config">
MathJax.Hub.Config({
  TeX: { equationNumbers: { autoNumber: "AMS" } }
});
</script>

@@

Chapter \@ref(kernel-inspection) gives the distribution of coefficients in the 
different kernel matrices used in hybrid prediction.

>>>>

@@

Chapter \@ref(original-kernels) gives the distribution of coefficients in the 
different kernel matrices used in hybrid prediction.

Chapter \@ref(transformed-kernels) gives the distribution of coefficients in 
$\mathbf{W}$ after applying centering, scaling, transformations and feature 
filtering on the original matrices.

### 2016-12-07 reports/single_step_kernels.Rmd
Describe kernel transformation in report

# Kernel Inspection
```{r setup-kernel-inspection, message = FALSE, echo = FALSE}
if (!require("pacman")) install.packages("pacman")
if (!require("devtools")) install.packages("devtools")
devtools::install_github("mwesthues/sspredr", update = TRUE)
pacman::p_load("BGLR","data.table", "tidyverse", "dtplyr", "caret", "e1071",
               "stringr")
pacman::p_load_gh("mwesthues/sspredr")

@@

Goal: Inspection of input matrices for single-step hybrid prediction using 
BGLR.
With the original configuration of the predictors, very low predictive abilities
were obtained for the amalgamation of pedigree, genomic and transcriptomic 
information, regardless of the phenotypic trait.
One of the possible causes were inconsistencies in the functions `impute_eta()`
and `impute2` from the `sspredr` package regarding the factorization of 
matrices.
After tackling those bugs, we wanted to revisit the distribution of the 
coefficients from the matrix $\mathbf{W}$ in \@ref(eq:mrna-model).

@@

```{r Plot-kernel-coefficients, cache = TRUE, fig.cap = "Distribution of coefficients that are elements of the kernel matrices used in single step or regular prediction of hybrid performance."}
## -- PLOT KERNEL COEFFICIENTS ---------------------------------------------
# Extract the coefficients from the lower triangle of the kernels for
# subsequent plotting of their distributions.
W_df <- W_lst %>%
  at_depth(.depth = 2, .f = ~.[match(colnames(.), rownames(.)), ]) %>%
  at_depth(.depth = 2, .f = ~.[lower.tri(., diag = FALSE)]) %>%
  map(function(x) {
    names(x) <- c("Dent", "Flint")
    x 
  }) %>%
  map(.f = stack) %>%
  bind_rows(.id = "id") %>%
  rename(Predictor = id, 
         Values = values,
         Group = ind) %>%
  as_tibble()

>>>>

# Original Kernels
```{r setup-kernel-inspection, message = FALSE, echo = FALSE}
if (!require("pacman")) install.packages("pacman")
if (!require("devtools")) install.packages("devtools")
devtools::install_github("mwesthues/sspredr", update = TRUE)
pacman::p_load("BGLR","data.table", "tidyverse", "dtplyr", "caret", "e1071",
               "stringr")
pacman::p_load_gh("mwesthues/sspredr")

@@

With the original configuration of the predictors, very low predictive abilities
were obtained for the amalgamation of pedigree, genomic and transcriptomic 
information, regardless of the phenotypic trait.
One of the possible causes were inconsistencies in the functions `impute_eta()`
and `impute2` from the `sspredr` package regarding the factorization of 
matrices.
After tackling those bugs, we wanted to revisit the distribution of the 
coefficients from the matrix $\mathbf{W}$ in \@ref(eq:mrna-model).

@@

```{r Plot-kernel-coefficients, cache = TRUE, fig.cap = "Distribution of coefficients that are elements of the unmodified kernel matrices ($\\mathbf{W}$) used in single step or regular prediction of hybrid performance."}
## -- PLOT KERNEL COEFFICIENTS ---------------------------------------------
# Extract the coefficients from the lower triangle of the kernels for
# subsequent plotting of their distributions.
W_df <- W_lst %>%
  at_depth(.depth = 2, .f = ~.[match(colnames(.), rownames(.)), ]) %>%
  at_depth(.depth = 2, .f = ~.[lower.tri(., diag = FALSE)]) %>%
  map(function(x) {
    names(x) <- c("Dent", "Flint")
    x 
  }) %>%
  map(.f = stack) %>%
  bind_rows(.id = "id") %>%
  rename(Predictor = id, 
         Values = values,
         Group = ind) %>%
  as_tibble()

### 2016-12-07 reports/transformed_kernels.Rmd
Describe kernel transformation in report

# Kernel Inspection
```{r setup-kernel-transformation, message = FALSE, echo = FALSE}
if (!require("pacman")) install.packages("pacman")
if (!require("devtools")) install.packages("devtools")
devtools::install_github("mwesthues/sspredr", update = TRUE)
pacman::p_load("BGLR","data.table", "tidyverse", "dtplyr", "caret", "e1071",
               "stringr")
pacman::p_load_gh("mwesthues/sspredr")

@@

Goal: Tranformation of predictor data for improved single-step prediction using 
BGLR.

@@

```{r Plot-transformed-kernel-coefficients, cache = TRUE, fig.cap = "Distribution of coefficients that are elements of the kernel matrices used in single step or regular prediction of hybrid performance."}
## -- PLOT KERNEL COEFFICIENTS ---------------------------------------------
# Extract the coefficients from the lower triangle of the kernels for
# subsequent plotting of their distributions.
W_df <- W_lst %>%
  at_depth(.depth = 2, .f = ~.[match(colnames(.), rownames(.)), ]) %>%
  at_depth(.depth = 2, .f = ~.[lower.tri(., diag = FALSE)]) %>%
  map(function(x) {
    names(x) <- c("Dent", "Flint")
    x 
  }) %>%
  map(.f = stack) %>%
  bind_rows(.id = "id") %>%
  rename(Predictor = id, 
         Values = values,
         Group = ind) %>%
  as_tibble()

>>>>

# Transformed Kernels
```{r setup-kernel-transformation, message = FALSE, echo = FALSE}
if (!require("pacman")) install.packages("pacman")
if (!require("devtools")) install.packages("devtools")
devtools::install_github("mwesthues/sspredr", update = TRUE)
pacman::p_load("BGLR","data.table", "tidyverse", "dtplyr", "caret", "e1071",
               "stringr")
pacman::p_load_gh("mwesthues/sspredr")

@@

After noticing extremely narrow distributions in the coefficients of 
$\mathbf{W}$ (defined in Eq. \@ref(eq:mrna-model), Fig.
\@ref(fig:Plot-kernel-coefficients)), we decided to explore the effect of several
data pre-processing procedures on the distribution of the coefficients.
Separately for each heterotic group, the following data pre-processing 
procedures were applied to the transcriptomic data: i) Box-Cox transformation,
ii) centering, iii) scaling, removal of features with a correlation 
$\rho \geq 0.999$, using the function `preProcess` from the R package `caret` 
[@Kuhn2013].
Separately for each heterotic group and each subset of features (i.e. `snp100`, 
`snp77` and `snp42`), only SNPs that were not in perfect linkage disequilibrium
with another marker were kept. 
From Fig. \@ref(fig:Plot-transformed-kernel-coefficients) we can glean that 
data processing had a considerable, positive effect on the distribution of 
coefficients in $\mathbf{W}$, which became wider in the case of predictor 
combinations such as `ped100_snp77_mrna42` compared to unprocessed features 
(Fig \@ref(fig:Plot-kernel-coefficients)).

@@

```{r Plot-transformed-kernel-coefficients, cache = TRUE, fig.cap = "Distribution of coefficients that are elements of the transformed kernel matrices $\\mathbf{W}$ used in single step or regular prediction of hybrid performance."}
## -- PLOT KERNEL COEFFICIENTS ---------------------------------------------
# Extract the coefficients from the lower triangle of the kernels for
# subsequent plotting of their distributions.
W_df <- W_lst %>%
  at_depth(.depth = 2, .f = ~.[match(colnames(.), rownames(.)), ]) %>%
  at_depth(.depth = 2, .f = ~.[lower.tri(., diag = FALSE)]) %>%
  map(function(x) {
    names(x) <- c("Dent", "Flint")
    x 
  }) %>%
  map(.f = stack) %>%
  bind_rows(.id = "id") %>%
  rename(Predictor = id, 
         Values = values,
         Group = ind) %>%
  as_tibble()
